---
title: "Assignment 4 - Applying meta-analytic priors"
author: "Riccardo Fusaroli"
date: "3/16/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 4

In this assignment we do the following:
- we reproduce the meta-analysis of pitch SD from last semester in a Bayesian framework
- we reproduce the pitch SD in schizophrenia analysis from last semester using both a conservative and a meta-analytic prior
- we assess the difference in model quality and estimates using the two priors.

The questions you need to answer are: What are the consequences of using a meta-analytic prior? Evaluate the models with conservative and meta-analytic priors. Discuss the effects on estimates. Discuss the effects on model quality. Discuss the role that meta-analytic priors should have in scientific practice. Should we systematically use them? Do they have drawbacks? Should we use them to complement more conservative approaches? How does the use of meta-analytic priors you suggest reflect the skeptical and cumulative nature of science?

### Step by step suggestions

Step 1: Reproduce the meta-analysis of pitch sd from previous studies of voice in schizophrenia
- the data is available as Assignment4MetaData.xlsx
- Effect size (cohen's d), sd and variance are already calculated (you're welcome!)
- Since we're only interested in getting a meta-analytic effect size, let's take a shortcut and use bromance magic (brms): https://mvuorre.github.io/post/2016/2016-09-29-bayesian-meta-analysis/

```{r, include=FALSE}}
library(brms)
library(brmstools)
library(dplyr)
#library(rethinking)
library(bayesplot)
library(readxl)

```


```{r}

dm = read_excel("Assignment4MetaData.xlsx")
dp = read_excel("Assignment4PitchDatav2.xlsx")


# cohens d is a standardized measure of effect size. It assumes a Gaussian distribution
# model estimates same effect size with 2000 and 10000 iterations.

# KØR model igen
m = brm(MeanES|se(SdES) ~ 1  + (1|StudyID),
        data = dm,
        cores = 2,
        chain = 2,
        iter = 10000,
        warmup = 4000
        )

summary(m)

pairs(m)

# plotting
plot(m)
forest(m,
       av_name = "Effect Size" )


```


Step 2: Prepare the pitch SD data from last year
- the data is available as Assignment4PitchData.csv (thanks Celine)
- We do not know how to build random effects, yet. So we need to simplify the dataset to avoid interdependence between datapoint: How?
- Also, let's standardize the data, so that they are compatible with our meta-analytic prior (Cohen's d is measured in SDs).

```{r}
# one data point per participant
# mean by participant, best way to summarize

dp2 = aggregate(dp[c(2,6,7)], list(dp$ID_unique), mean)

dp2 = rename(dp2, replace = c("Group.1" = "ID_unique"))

# scaling
dp2$PitchMean = scale(dp2$PitchMean)
dp2$PitchSD = scale(dp2$PitchSD)

```


Step 3: Build a regression model predicting Pitch SD from Diagnosis.
- how is the outcome distributed? (likelihood function)
- how are the parameters of the likelihood distribution distributed? Which predictors should they be conditioned on?
- use a skeptical/conservative prior for the effects of diagnosis. Remember you'll need to motivate it.
- Describe and plot the estimates. Evaluate model quality

```{r}

# model
# when the variable is continious we can assume that it is normally distributed as most phenomena are (least error)
# central limit theorem, when the variable is continious, the distribution of the mean of the distribution tends to be Gaussian

'''
PitchSD ~ Normal(mu,sigma)
mu = a+ b*diagnosis
a ~ Normal(0,1) # (0,2) very broad
b ~ Normal(0,0.1) # we will get back to that
sigma ~ Cauchy(0,2) # Uniform, terrible for sampling

'''
# map 2 stan
m3 <- map2stan(
  alist(
    PitchSD ~ dnorm(mu,sigma) ,
    mu <- a + b*diagnosis ,
    a ~ dnorm(0,1),
    b ~ dnorm(0,0.1),
    sigma ~ dcauchy(0,2)
    ), 
  data= dp2)

summary(m3)


# BRMS
# check which priors can be set
get_prior(PitchSD ~ diagnosis, data = dp2)
m3 = brm(PitchSD ~ diagnosis,
         prior = c(prior(normal(0,1), class = Intercept),
         prior(normal(0,0.1), class = b, coef = diagnosis),
         prior(cauchy(0,2), class = sigma)),
         data = dp2,
         cores = 4,
         chain = 4,
         iter = 20000,
         warmup = 4000
         )

summary(m3)
plot(m3)

```


Step 4: Now re-run the model with the meta-analytic prior
- Describe and plot the estimates. Evaluate model quality

```{r}

# map2stan
# if you choose se as sigma of prior distribution, you care about the underlying true effect
# if you choose SD as sigma of prior distribution, you care about the thruth of your study.
# depends on what you think about the existing litterature and what question you are asking.

# the below model gives the following error: Error in constr_list[[sigma_name]] : no such index at level 1
m4 <- map2stan(
  alist(
    PitchSD ~ dnorm(mu,sigma) ,
    mu <- a + b*diagnosis ,
    a ~ dnorm(0,1),
    b ~ dnorm(0.6,0.3),
    sigma ~ dcauchy(0,2)
    ), 
  data= dp2)


# BRMS
# check which priors can be set
get_prior(PitchSD ~ diagnosis, data = dp2)


m4 = brm(PitchSD ~ diagnosis,
         prior = c(prior(normal(0,1), class = Intercept),
         prior(normal(0.6,0.3), class = b, coef = diagnosis),
         prior(cauchy(0,2), class = sigma)),
         data = dp2,
         cores = 4,
         chain = 4,
         iter = 50000,
         warmup = 10000
         )

summary(m4)
plot(m4)

# NULL model
m5 = brm(PitchSD ~ 1,
         prior = c(prior(normal(0,1), class = Intercept),
         prior(cauchy(0,2), class = sigma)),
         data = dp2,
         cores = 4,
         chain = 4,
         iter = 50000,
         warmup = 10000
         )

summary(m5)
plot(m5)

```


Step 5: Compare the models
- Plot priors and posteriors of the diagnosis effect in both models
- Compare posteriors between the two models
- Compare their relative distance from truth (WAIC)
- Discuss how they compare and whether any of them is best.
```{r}

# plotting priors

x <- seq(from = -1, to = 2, by = 0.01)
prior_c = dnorm(x,mean = 0, sd = 0.1)
prior_m = dnorm(x,mean = 0.6, sd = 0.3)

plot(x, prior_c, type = "l", col = "red")
lines(x, prior_m, type = "l", col = "green")


# plotting posteriors
plot(m3)
plot(m4)

plot(marginal_effects(m3), ask = FALSE)
plot(marginal_effects(m4), ask = FALSE)


pp_check(m3, nsamples = 50)
pp_check(m4, nsamples = 50)
pp_check(m5, nsamples = 50)


# compare models
waic(m3, m4, m5) #  waic does not work on brm models when rethinking package is loaded

waic_m3 = waic(m3)
waic_m4 = waic(m4)
waic_m5 = waic(m5)

# load rethinking package
compare(waic_m3, waic_m4, waic_m5) 



```


Step 6: Prepare a nice write up of the analysis and answer the questions at the top.

Optional step 7: how skeptical should a prior be?
- Try different levels of skepticism and compare them using WAIC.

Optional step 8: Include other predictors
- Do age, gender and education improve the model?
- Should they be main effects or interactions?

Optional step 9: Bromance magic.
- explore the bromance code below including random effects (by default with weakly informative priors)
- learn how to change the prior
- explore effects of trial, age, gender, including the appropriate random slopes
- compare the models you created using WAIC and posterior predictive check (pp_check())


```{r}

brm_out <- brm(PitchSD ~ 1 + Diagnosis  +(1|ID_unique/Study), # Outcome as a function of the predictors as in lme4. 
               data=Data, # Define the data
               family=gaussian(), # Define the family. 
               iter = 5000, warmup = 2000, cores = 4)
summary(brm_out1)
plot(brm_out1)

```

